\thispagestyle{empty}

\begin{center}
    {\LARGE\bf Abstract}
\end{center}

FundFind is a web application which enables scholars, research development officers and postgraduates to share information about funding opportunities. Data about funding opportunities and funding organisations can be submitted, browsed and searched via a mobile-friendly web UI as well as a JSON API. FundFind communicates with its data store via simple asynchronous HTTP. This enabled the development of a secondary Funding Harvest application which harvests funding data automatically and is completely separate from FundFind - it simply puts data into the same datastore.

The project is a learning exercise, aiming to explore the field of scholarly funding, identifying users and their needs, as well as potential problems with the open sharing of such data. These insights and the technical outputs are meant to be developed further (by an interested party) into a funding information discovery service after this project is completed. The main application was based on another prototype service developed by the same party.

The project did allow for a lot of insight into the funding sector and users of funding information. However, the parent project had no automated testing and difficulties with integrating multiple third-party pieces of software hampered technical development. The incomplete technical prototypes impeded the depth of the insights gained as feasible user testing was limited. The main application is, however, of similar quality to its parent prototype. The lessons learned are sufficient and valuable for the future development of a FundFind service.